---
title: "Time Series Project 1 "
author: "PUT YOUR NAME HERE"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Project Description

In this project, you are going to test your skills that you've learned so far in the course on a few classical time series datasets. For each dataset, I'll give you a set of directions that I'd like you complete. Make sure that you answer each question fully and comment on the outcomes when appropriate. 



## Lynx Data

For this part of the project, we are going to work with data on lynx. Specifically, we are looking at the annual number of lynx trapped in the Mackenzie River district of northwest Canada. The dataset resides in the `datasets` package and is called `lynx`. Complete the following tasks:

  * The dataset should already be loaded into your R workspace by default, but if it isn't, run `data("lynx")` to load the data.
  * Plot the data and comment on any major features that you see in the data (trend, periodocity, stationarity, shape, white-noise or not, etc.)
  * Plot the correlogram and comment on the major trends you see.
  * Due to the asymmetrical peaks and smooth troughs of the series, it will better to transform the data using the log base 10 transformation. In this, I mean take the log base 10 of the series using the `log10()` function. Plot the new series as well as the correlogram and comment on the differences and similarities with the original series.
  * In a paper by Tong in 1977, the lynx data was fit with an AR(11) model. I want you to fit an AR(11) model to the data using the maximum likelihood method. Specifically, you want to use the `ar()` function like we are used to doing. Luckily, the best model is chosen to be an AR model of order 11 by default. Also, assess how the model fit. In addition to the usual methods, we can add a new tool to our toolbox. The Ljung-Box test is a statistical test to assess whether or not a series is white-noise. If the p-value is less than a nominal 0.05 level, then it is not white noise and if it is greater than 0.05, then it can't be concluded that it's not white-noise. You'll want to run the following code `Box.test(data, lag = 24, type = "Ljung-Box")`
  * Another model that has been suggested is an ARMA(2,3). Fit an ARMA(2,3) model to the data and assess the fit.
  * Compare and contrast the two model fits.
```{r}
lynx %>% plot()
lynx %>% acf()
```

The data appears to be lacking in overall trend with a seasonal trend that appears to go up to a peak, then fall back down to around 200 every 10 or so years. This trend seems to vary with a larger peak happening every 4th time and smaller peaks occurring the other three times in the cycle. This data appears to be lacking in stationary as there is an seasonal trend and also looking at the corellorgram there are statistically significant correlations negative around 5 and 15 years, and positively around 10 and 20 years. 
```{r}
lynx %>% log10() %>% plot()
lynx %>% log10() %>% acf()
```

The new series seems to have the same amount of peaks and valleys as the original series, however instead of starting low and rising up, this series has more of the shape of a sin wave where it starts in the middle and goes up and down. The seasonal pattern seems to have remained but the cycle extreme peaks seem to have disappeared. The correlogram looks almost the same, some of the correlations are stronger than before, however they all say the same thing as before. 



```{r}
tseries::adf.test(lynx %>% log10)

thing1 <- arima(lynx %>% log10(), c(11, 0, 0)) 

thing1 %>% summary()

Box.test(thing1$resid, lag = 24, type = "Ljung-Box")



thing1$resid %>% plot()
thing1$resid[12:114] %>% acf()

thing1 %>% AIC()

```



```{r}
tseries::adf.test(lynx %>% log10)

thing2 <- arima(lynx %>% log10(), c(2, 0, 3))

thing2 %>% summary()

Box.test(thing2$resid, lag = 24, type = "Ljung-Box")



thing2$residuals %>% plot()
thing2$residuals[12:114] %>% acf()

thing2 %>% AIC()

```


## Wages Data

This next dataset we are going to work with is the wages data. Specifically, the dataset contains the average English daily wages in pounds for each year from 1260 to 1944. This dataset is from the `tswge` package. You will need to install the package first. After installing and loading the package, run `data("wages")` to load the data into the workspace. Complete the following tasks:

  * Plot the realization and comment on what you see (just like the first dataset)
  * Compute the first-order difference of the series using the `diff()` function and plot the differenced data. Then plot the correlogram and partial acf plot. Comment on if the differenced series is white noise. 
  * Fit an ARMA(2,1) and assess the fit. 
  * Fit an ARIMA(2,1,1) and assess the fit.
  
  

## Choose Your Own

In this part of the project, I want you to pick a dataset of your choosing and perform some analysis! I'll leave the actual analysis choices up to you, but you are required to run your ideas past me first. 















